\MOC{} Server Pages:

### 2007-04-16 paper/paper.tex
svn-r18: Changes to support re-executing stage 1 once server is live.
darcs-hash:20070416222509-e5a07-6b6a7f694ec557c1730173da3206b03b08dd89da.gz

\screenshot{confcal}{A simple conference calendar,
  implemented as a PHP script that connects to a PostgreSQL
  database.  Not as sophisticated as the service described
  above!}

@@

The general idea of a \emph{server page} language is that we
  write HTML by default, and embed code %
\lstinline{<?like this?>}.  %
This is the standard XML syntax for \emph{processing
instructions.}  In our case, the instructions are written in
\MOC{}.   Here is your first \MOC{} server page:

@@

To explore the expressiveness of this design, we now look at a
series of web services, organized to comprise a small web
site.\footnote{I hope to make the demo available on the web in time
  for the workshop; this will require convincing the I.T. folks to
  open up a non-standard port through the firewall.  For now, it may
  be downloaded and run locally.} %
To give the services a similar look, we developed a site-wide style
sheet, and an OCaml function to generate a navigation bar from a
hierarchical list of page titles and links.  The follow bindings are
available site-wide:
\begin{lstlisting}[numbers=none]
  val preamble: string -> string
  val navbar: string -> string
  val postamble: string
\end{lstlisting}
The first function generates the standard XHTML header given the
title, while the second function generates the navigation bar given
the URI denoting the current page.  These can be called at either
stage.  Naturally, if the title includes a dynamic argument, then
\cd{preamble} will be delayed until the serve stage.

@@

The first function wraps a string into a fixed width block.  The power
function itself begins on line~\ref{line:power-fun}.  No output should

occur until the preamble is generated on
line~\ref{line:power-preamble}.  The user's input is translated to a
number relatively late (line~\ref{line:power-num}); if there is a
problem here, the navigation bar and form will have already been
output before the exception message.

@@

\textit{Ugh\ldots out of time\ldots but I will explain one
  critical detail\ldots}

@@

\citet{elsman04typing}, \citet{wallace99haxml},
\citet{hosoya03xduce}, and \citet{ohl04xhtml} validate HTML code
in functional languages.  We could probably use OCaml's type
system to greater advantage this way, but it may require
validating the text parts of the server page with a more
mainstream XML parser.

>>>>

\screenshot{confcal}{A simple conference calendar, implemented as a
  PHP script that connects to a PostgreSQL database.  Not as
  sophisticated as the service described in the text.}

@@

The general idea of a \emph{server page} language is that we
  write HTML by default, and embed code %
\lstinline{<?like this?>}.  %
PHP programmers are familiar with this syntax for embedding code, but
in our case, the code itself is written in \MOC.  Here is your first
\MOC{} server page:

@@

To explore the expressiveness of this design, we now look at a
series of web services, organized to comprise a small web
site.\footnote{I hope to make the demo available on the web in time
  for the workshop; this will require convincing the I.T. folks to
  open up a non-standard port through the firewall.  For now, it may
  be downloaded and run locally.} %
To give the services a similar look, we developed a site-wide style
sheet, and an OCaml function to generate a navigation bar from a
hierarchical list of page titles and links.  The following bindings are
available site-wide:
\begin{lstlisting}[numbers=none]
  val preamble: string -> string
  val navbar: string -> string
  val postamble: string
\end{lstlisting}
The first function generates the standard XHTML header given the
title, while the second function generates the navigation bar given
the URI denoting the current page.  These can be called at either
stage.  Naturally, if the title includes a dynamic argument, then
\cd{preamble} will be delayed until the serve stage.

@@

The first function wraps a string into a fixed width block.  The power
function itself begins on line~\ref{line:power-fun}.  No output should
occur until the preamble is generated on
line~\ref{line:power-preamble}.  The user's input is translated to a
number relatively late (line~\ref{line:power-num}); if there is a
problem here, the navigation bar and form will have already been
output before the exception message.

@@

@@

\citet{elsman04typing}, \citet{wallace99haxml},
\citet{hosoya03xduce}, and \citet{ohl04xhtml} validate HTML code
in functional languages.  We could probably use OCaml's type
system to greater advantage this way, but it may require
validating the text parts of the server page with a more
mainstream SGML or XML validator.
